---
title: "Parole"
output: github_document
editor_options: 
  chunk_output_type: console
---
## Load libraries
```{r libraries}
options(scipen = 2)
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(googlesheets4)
```


## Roster

### Read
Inmate roster contains everyone who's ever involved in the Nebraska prison system.
```{r roster read}
inmateDB_path <- here("parole","inmateDB","inmateDB1.csv")
inmateDB <- read_csv(inmateDB_path)
inmateDB2_path <- here("parole","inmateDB","inmateDB2.csv")                      
inmateDB2 <- read_csv(inmateDB2_path)
inmateDB <- inmateDB %>% clean_names()
inmateDB2 <- inmateDB2 %>% clean_names()
```
### Clean columns
```{r roster clean}
inmateDB1 <- inmateDB %>% 
  mutate_at(.vars = vars(ends_with("date")), .funs = as.Date, format = "%m/%d/%Y")
#inmateDB1$id_number = as.character(inmateDB1$id_number)
remove(inmateDB)
```
### Layout
```{r roster layout}
inmateDB1 <- inmateDB1 %>% 
  select(30,16,17,27,1:3,10:12,everything())
```

## Parole by votes 

### Read
clean columns
```{r read parole by vote}
pr_votes_path <- here("parole","parole_by_votes.csv")
pr_votes <- read_csv(pr_votes_path)
pr_votes <- pr_votes %>% clean_names()
pr_votes1 <- pr_votes %>%
  mutate(hearing_date = as.Date(hearing_date, format = "%m/%d/%Y"))
```
### Fill down rows
Writing a function here to fill down the motion for all five members. 
```{r votes data fill results}
pr_votes1 <- pr_votes1 %>% 
  fill(1:3,.direction = "down")
# pr_votes1 %>% group_by(id_number, hearing_date) %>% summarise(n=n()) %>% .$n %>% table()
fill_results <- function(df) {
  for (i in 0: (nrow(df)/5-1)) {
    df_cols = which(!is.na(df[5*i + 1,4:7]) %>% as.vector())
    df_cols = df_cols + 3
    for (df_col in df_cols) {
      if (!is.na(df[5*i+1,df_col])) {
            df[i*5+2,df_col] = df[5*i+1,df_col]
            df[i*5+3,df_col] = df[5*i+1,df_col]
            df[i*5+4,df_col] = df[5*i+1,df_col]
            df[i*5+5,df_col] = df[5*i+1,df_col]
      }
    }
  }
  return(df)
}

pr_by_votes <- fill_results(pr_votes1)
#decisions_number <- rowSums(is.na(pr_by_votes))
#pr_val <- pr_by_votes %>% add_column(decisions_number)
```
### Summarize
#### individual by board member
```{r votes wider individual level results}
pr_ind <- pr_by_votes %>% 
  #group_by(id_number, inmate_name, hearing_date,vote) %>% summarize(count  = n())
  pivot_wider(names_from = board_member_last_name, values_from = vote) 
# new table by motion, member, vote
pr_motion <- pr_by_votes %>% 
  pivot_longer(cols = c(paroled,denied,deferred),names_to = "motion") 
pr_motion <- pr_motion %>% 
  filter(!is.na(value)) %>% select(-value)
```
#### individual vote pattern

```{r individual pattern}
# new table by motion and type of vote counts
pr_ind_pattern <- pr_motion %>% group_by(id_number, hearing_date,vote,motion) %>% summarize(number = n())

pr_ind_pattern <- pr_ind_pattern %>% pivot_wider(names_from = vote, values_from = number) %>% clean_names()
```

### Compare missing
Of these `r nrow(pr_ind_pattern)` hearings, `r sum(pr_ind_pattern$missing_member == "MISSING MEMBER")`
```{r compare missing}
# add parole indicator column 
pr_ind_pattern <- pr_ind_pattern %>% 
  mutate(is_paroled = if_else(condition = motion == "paroled" & yes >=3, true = "PAROLED", false = "NOT PAROLED"),
         missing_member = if_else(is.na(not_available), true = "NO MISSING", false = "MISSING MEMBER"))

with(pr_ind_pattern, table(missing_member, is_paroled))

# when at least a board member was missing, the parole board granted parole 56.2% of the time, versus 62.6% when all five members voted. 
# when there were missing members
pr_ind_pattern %>% filter(not_available >= 1) %>% .$motion %>% tabyl()
pr_ind_pattern %>% filter(not_available >= 1) %>% .$is_paroled %>% tabyl()

# without missing members
pr_ind_pattern %>% filter(is.na(not_available)) %>% .$motion %>% tabyl()
pr_ind_pattern %>% filter(is.na(not_available)) %>% .$is_paroled %>% tabyl()


# There were 2,491 hearings with one missing member. There were two hearings with 2-2 votes but those were deferment anyway.
split_votes <- pr_ind_pattern %>% filter(not_available == 1) %>% arrange(desc(no )) 
split_votes$is_paroled %>% tabyl()
```

#### Race
```{r missing racial pattern}

#add race description
pr_ind_pattern1 <- pr_ind_pattern %>% 
  left_join(inmateDB1 %>% select(id_number, race_desc), by = "id_number")
#missing status by race by motion
pr_ind_pattern2 <- pr_ind_pattern1 %>% group_by(race_desc, motion, missing_member) %>% summarize(n = n()) %>% left_join(pr_ind_pattern1 %>% group_by(missing_member,race_desc) %>% summarize(total_hearings =n()))

pr_ind_pattern2 <- pr_ind_pattern2 %>% mutate(motion_rate = n/total_hearings)

pr_ind_pattern2 %>% ggplot(aes(y = motion_rate, x = race_desc, fill = missing_member)) +
  geom_bar(position = "dodge", stat = "identity") + facet_wrap(~motion,nrow = 3)
```

### Board members' voting patterns
```{r votes data by member}
pr_bm <- pr_motion %>% group_by(board_member_last_name, motion,vote) %>% summarize( count = n())
pr_bm <- pr_bm %>% add_tally(count)  
pr_bm <- pr_bm %>% mutate(type_rate = count/n)

library(wesanderson)
pr_bm %>% ggplot(aes(x=motion,y=count, fill = vote)) + geom_col() +
  facet_wrap(~board_member_last_name) +
  scale_fill_manual(values=wes_palette(n=5, name="Darjeeling1"))+
  theme_minimal()
```

