---
title: "Parole hearing data diary"
output: github_document
editor_options: 
  chunk_output_type: console
---
## Load libraries
```{r libraries,results='hide'}
options(scipen = 2)
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(campfin)
library(calendar)
library(googlesheets4)
```

## Parole by votes 
The data was obtained via a public records request filled by the Nebraska Board of Parole. We ask three data questions: 
1. How many individual cases/hearings had at least one absence and how many had the full board in attendance?
2. How many full days had at least one board member missing the full day, what's the percentage of absence days out of all hearing days?
3. What's the parole grant rate when the parole was full, short 1,2,and 3 members respectively? 
### Read
clean columns
```{r read parole by vote}
pr_votes_path <- here("parole","parole_by_votes.csv")
pr_votes <- read_csv(pr_votes_path)
pr_votes <- pr_votes %>% clean_names()
pr_votes1 <- pr_votes %>%
  mutate(hearing_date = as.Date(hearing_date, format = "%m/%d/%Y"))
```

### Fill down rows
In the original data, each individual hearing takes up five lines, with the first line showing the motions (there are seveal columns `deferred`,`paroled`,`denied` and `defer to md`, which means deferred to mandatory discharge and is not a motion of its own but often accompanies a denial). 

First, I filled down the person's id number, name and hearing days. Then I wrote a function `fill_results()` here to fill down the motion for all five members. 

```{r votes data fill results}
pr_votes1 <- pr_votes1 %>% 
  fill(1:3,.direction = "down")
# pr_votes1 %>% group_by(id_number, hearing_date) %>% summarise(n=n()) %>% .$n %>% table()
fill_results <- function(df) {
  for (i in 0: (nrow(df)/5-1)) {
    df_cols = which(!is.na(df[5*i + 1,4:7]) %>% as.vector())
    df_cols = df_cols + 3
    for (df_col in df_cols) {
      if (!is.na(df[5*i+1,df_col])) {
            df[i*5+2,df_col] = df[5*i+1,df_col]
            df[i*5+3,df_col] = df[5*i+1,df_col]
            df[i*5+4,df_col] = df[5*i+1,df_col]
            df[i*5+5,df_col] = df[5*i+1,df_col]
      }
    }
  }
  return(df)
}

pr_by_votes <- fill_results(pr_votes1)
```
### Summarize
#### Data manipulation
pr_ind: transpose the data from (board member last name and their votes) into (cotton, gissler, etc.) still 6521 individual hearings
pr_motion: collapse the three columns `paroled`, `denied` and `deferred` into one single column motion
Then since we'll end up with lots of empty rows resulted from votes on decisions on motions that weren't actually made. It has 97815 rows. 
Then we need to get rid of empty rows. By doing this we also exclude hearings with no motions indicated in the data. Now we have 6,643 actual hearings.
```{r votes wider individual level results}
pr_ind <- pr_by_votes %>% 
  pivot_wider(names_from = board_member_last_name, values_from = vote) 
# new table by motion, member, vote
pr_motion <- pr_by_votes %>% 
  pivot_longer(cols = c(paroled,denied,deferred),names_to = "motion") 
pr_motion <- pr_motion %>% 
  filter(!is.na(value)) %>% select(-value)
```
#### individual vote pattern

```{r individual pattern}
# new table by motion and type of vote counts, 36150 records 
pr_ind_pattern <- pr_motion %>% group_by(id_number,hearing_date,vote,motion) %>% summarize(number = n())
# pivot from vote-yes,not available, no to how many yes votes, how many no votes and how many nas. We 
pr_ind_pattern <- pr_ind_pattern %>% pivot_wider(names_from = vote, values_from = number) %>% clean_names()
```


### Compare missing
Of these `r nrow(pr_ind_pattern)` hearings, `r sum(pr_ind_pattern$missing_member == "MISSING MEMBER")` had missing members.
```{r compare missing}
# add parole indicator column & absence indicator column
pr_ind_pattern <- pr_ind_pattern %>% 
  mutate(is_paroled = if_else(condition = motion == "paroled" & yes >=3, true = "PAROLED", false = "NOT PAROLED"),
         missing_member = if_else(is.na(not_available), true = "NO MISSING", false = "MISSING MEMBER"))
```

Number of hearings attended by partial v. full board 
This answers the first question: How many individual hearings had the full board in attendance and how many had at least one member missing?
```{r missing vs. fully present}
pr_ind_pattern$missing_member %>% tabyl()
```

### Number of hearing days with at least one absence
This answers the second question: 2. How many full days had at least one board member missing the full day, what's the percentage of absence days out of all hearing days?
```{r}
#all the types of votes a board member ever cast that day. how many were yes, no, not available
pr_miss <- pr_motion %>% group_by(board_member_last_name, hearing_date,vote) %>% summarize( count = n())
#the total number of votes that day (should be equal to the total of yes, no and not available votes)
pr_miss_tj <- pr_miss %>% group_by(board_member_last_name, hearing_date) %>% summarize( total = sum(count))
# join them together, so we see what portion of the vote that day was yes, no or not_available
pr_miss <-  pr_miss %>% left_join(pr_miss_tj)
# calculate the rate 
pr_miss <- pr_miss %>% mutate(missing_rate = count/total)

# meetings where board members missed the full day of meetings, but if one day two board members were absent that date would show up twice -- we have 253 records
pr_miss_true <- pr_miss %>% filter(missing_rate ==1 & vote == "Not Available")

# how many board members missed that full day by date. The actual number of days is 184.
# 119 days with one board member absent, 69 days when two board members were missing
pr_board_miss <- pr_miss_true %>% group_by(hearing_date) %>% summarize(member_absent = n())

# how many days did the board have hearings, in total, count the distinct days -- 322
pr_ind_pattern$hearing_date %>% n_distinct()
# it's the same thing if we apply it to the original data
pr_votes1$hearing_date %>% n_distinct()
```

184/322 = 57%


#### Parole grant rate in cases of a 5-member board, 4-member board and 3-member board.

```{r}

# look at how many assessments were approved and total hearings heard by 5,4,3,and2 board members
#here I also substituted all the NAs (not_available was NA means no one was unavailable)
cross_tabs <- xtabs(~not_available + is_paroled, data = pr_ind_pattern %>% mutate(not_available = replace_na(not_available, replace = 0)))

cross_tabs
# cross tabulation in percentage
100* prop.table(cross_tabs, margin = 1) -> cross_tabs_perc
cross_tabs_perc
```

